\section{Experiments}
\label{sec:experiments}

To evaluate the effectiveness of the proposed adaptive sampling methods, we conduct experiments across a range of challenging two-dimensional PDE problems. Our experimental framework, \textbf{\pinnpoint}, is designed to ensure a rigorous and fair comparison between sampling strategies.

\subsection{Benchmark Problems}

We employ a suite of four distinct PDE problems, each presenting unique challenges for PINN training:

\begin{enumerate}
    \item \textbf{Poisson Equation on Perforated Domain:} A steady-state elliptic problem $-\nabla^2 u = f$ defined on a square domain with complex circular and cross-shaped cutouts. This problem tests the ability of the sampler to handle non-trivial geometries and boundary conditions.
    \item \textbf{Advection-Diffusion:} A problem characterized by high P\'{e}clet numbers, where the solution exhibits sharp fronts and localized features that are difficult to capture with uniform sampling.
    \item \textbf{Navier-Stokes (Channel with Obstacle):} A coupled system of equations representing fluid flow. The presence of an obstacle in the channel creates complex wake patterns and high-residual regions downstream.
    \item \textbf{Allen-Cahn Equation:} A reaction-diffusion equation used to model phase separation. It features sharp interfaces that evolve over the domain, requiring the sampler to dynamically track moving features (in our pseudo-time or iterative setup).
\end{enumerate}

\subsection{Fair Comparison Protocol}

A key contribution of our work is the establishment of a fair benchmarking protocol. All methods compared in this study share the following constraints:

\begin{itemize}
    \item \textbf{Shared Supervision:} Every method utilizes the same fixed set of labeled data points derived from a high-fidelity Finite Element Method (FEM) reference solution.
    \item \textbf{Fixed Iteration Budget:} All methods are trained for the same number of iterations and optimizer steps, ensuring that no method benefits from additional compute during the training phase.
    \item \textbf{Matched Collocation Budget:} The number of interior collocation points evaluated per iteration is kept constant across all methods.
    \item \textbf{Overhead Reporting:} We explicitly record and report the computational overhead associated with the sampling and mesh-refinement logic, allowing for a true error-to-cost evaluation.
\end{itemize}

\subsection{Experimental Setup}

The PINN models are implemented using PyTorch and consist of feed-forward neural networks with Swish activation functions. We use the Adam optimizer for the initial training phase, followed by L-BFGS for fine-tuning where appropriate. All experiments are run with multiple seeds (typically 10) to ensure statistical significance, with results reported as mean values with standard deviation shaded regions.
